---
title: "qbs121_hw8_gibran"
author: "Gibran Erlangga"
date: "3/9/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1 Data Analysis: Propensity Scores

Use the dataset called “Teaching Hospital Outcome.txt”. The outcome of interest is unfavorable discharge (discharge of a patient that is not to their home but to a nursing facility). The exposure of interest is whether the patient is in a teaching hospital or not. The covariates (potential confounders) are age, sex, race (white=referent, black, other), hispanic ethnicity, diabetes and hypertension.
``` {r}
data <- read.delim("Teaching Hospital Outcomes.txt")
head(data, 3)
```

1. Estimate the odds ratio between unfavorable discharge and teaching hospital controlling for the other variables provided (e.g. potential confounding variables) using a logistic regression model.
``` {r}
logit <- glm(UNFAVDX ~ ., family=binomial, data=data)
summary(logit)

# odds ratio
print('OddsRatio')
exp(logit$coefficients)
```

2. Develop a prediction model for whether or not patients received care at a teaching hospital using the method of random Forests. Use this model to calculate the propensity of receiving care at a teaching hospital.
``` {r}
library(randomForest)
library(tidyverse)
library(janitor)

data %>%
  clean_names() -> data_clean

# calculate propensity of TeachingHospital
rf <- randomForest(as.factor(teaching_hospital) ~ ., data=data_clean)
rf

votes <- rf$votes
estimated <- votes[,2]
estimated %>% summary()
```
3. Create 10 bins of the propensity score (e.g deciles of the propensity scores). Estimate the odds ratio between unfavorable discharge and teaching hospital controlling for this categorical bin variable. Note: The random forest approach may not yield propensities with a alot of distinct values, which may lead to an error when the bins are created.
``` {r}
Bin = cut(estimated, unique(quantile(estimated, (0:10)/10, na.rm=TRUE)))

model_binned <- glm(unfavdx ~. + Bin , data=data_clean, family = binomial)
summary(model_binned)

exp(model_binned$coefficients)
```

4. Calculate inverse propensity weights (after trimming the propensities so none is smaller than 0.01). Estimate the odds ratio between unfavorable discharge and teaching hospital using these weights.
``` {r}
# clip lower and upper values
estimated_norm <- ifelse(estimated < 0.1, 0.1, estimated)
estimated_norm <- ifelse(estimated > 0.9, 0.9, estimated)

inverse_prop_weights <- ifelse(data_clean$unfavdx, 1/estimated_norm, 1/(1-estimated_norm))
summary(inverse_prop_weights)

glm <- glm(as.factor(unfavdx) ~ ., family=binomial, weights = inverse_prop_weights, data=data_clean)
summary(glm)

exp(glm$coefficients)
```