---
title: "qbs121_hw7_gibran"
author: "Gibran Erlangga"
date: "2/28/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Analyses
## 2.1 PBC (50 points)
Use the dataset pbc that comes with the library survival in R. The time and status indicators compose the endpoint.
``` {r}
library(survival)
head(pbc)
```

1. Competing Risks: The variable status has 3 levels, 0 for censoring, 1 for transplant, 2 for death.
(a) (2pts) If someone had a transplant and died sometime later what would this data record? 

The status should show 2 for someone who had transplant and died when the study is running.

(b) Download the library, cmprsk
``` {r}
library(cmprsk)
```

(c) (3) Run the command to get “cumuative incidence” o=with(pbc, cuminc(time,status))
``` {r}
o <- with(pbc, cuminc(time, status))
o
```

(d) (5) Plot the output, plot(o), and interpret the resulting curves
``` {r}
plot(o)
```

2. (2) What proportion of patients are censored?
``` {r}
library(tidyverse)

pbc %>%
group_by(pbc$status == 0) %>%
summarise(n=n()) %>%
mutate(freq = n/sum(n))
```

3. (3) What is the total follow-up time in this study?
``` {r}
sum(pbc$time)
```

4. (5) Calculate the incidence rate of transplant and death (e.g. number of events divided by total follow-up time)
``` {r}
sum(pbc$status == 1)/sum(pbc$time)
```
5. Evaluate if trt arm impacts the composite endpoint, transplant or death by

(a) (2) first coding the composite endpoint, tr.death=status> 0
``` {r}
tr.death <- pbc$status > 0
```

(b) (3) reporting the number of events and rate in each arm
``` {r}
tapply(tr.death, pbc$trt, sum)
```

(c) (5) plotting survival curves and reporting median time to transplant or death in each treatment group
``` {r}
plot(survfit(Surv(time/3625.5, tr.death) ~ trt, data = pbc),
     fun = "event", 
     col=7:8,
     xlab="Years")

legend(1,0.7,
       text.col=7:8,
       legend=c("Drug","Placebo"))
```
(d) (5) conducting a log-rank test
``` {r}
survdiff(Surv(time, tr.death) ~ trt, data = pbc)
```
(e) (5) conducting a Peto-Peto-Prentice test (generalization of the Wilcoxon rank test)
``` {r}
survdiff(Surv(time, tr.death) ~ trt, rho=1, data=pbc)
```
(f) (5) using a Weibull proportional hazards model model to estimate the hazard ratio with a 95%CI
``` {r}
summary(survreg(Surv(time, tr.death) ~ trt, dist="weibull",data=pbc))
```
(g) (5) using a Cox’s proportional hazards model to estimate the hazard ratio with a 95%CI
``` {r}
summary(coxph(Surv(time, tr.death) ~ trt, data = pbc))
```

## 2.2 Lung
Use the dataset lung that comes with the library survival in R. The time and status indicators compose the endpoint.

1. (2pts) What are the levels of the status variable?
``` {r}
unique(lung$status)
```
1=censored, 2=dead

2. (2) What proportion of patients are censored?
``` {r}
perct_censored <- lung %>%
group_by(lung$status == 1) %>%
summarise(n=n()) %>%
mutate(freq = n/sum(n))

as.numeric(perct_censored[2,3]*100)
```

3. (3) What is the total follow-up time in this study?
``` {r}
sum(lung$time)
```

4. (3) Calculate the mortality rate (e.g. number of deaths divided by total follow-up time)
``` {r}
sum(lung$status == 2) / sum(lung$time)
```

5. Examine the association of biological sex with survival as follows
(a) (3) plotting male and female survival curves
``` {r}
surv_gender <- survfit(Surv(time, status) ~ sex, data = lung)
plot(surv_gender)
```

(b) (3) reporting median survival time for each biological sex
``` {r}
male <- lung %>%
filter(lung$sex == 1) #male

female <- lung %>%
filter(lung$sex == 2) #female

print(median(male$time))
print(median(female$time))
```

(c) (3) conducting a log-rank test
``` {r}
survdiff(Surv(time, status) ~ sex, data = lung)
```

(d) (3) conducting a Peto-Peto-Prentice test (generalization of the Wilcoxon rank test)
``` {r}
survdiff(Surv(time, status) ~ sex, rho=1, data = lung)
```

(e) (5) using a Weibull proportional hazards model model to estimate the hazard ratio with a 95% CI
``` {r}
summary(survreg(Surv(time, status) ~ sex, dist="weibull", data = lung))
```

(f) (5) using a Cox’s proportional hazards model to estimate the hazard ratio with a 95% CI
``` {r}
summary(coxph(Surv(time, status) ~ sex, data = lung))
```

6. (2) Combine categories 2 and 3 in variable ph.ecog and report survival curves by this new variable
``` {r}
lung$ph.ecog.females <- factor(ifelse(lung$ph.ecog >2, 2, lung$ph.ecog))
surv <- survfit(coxph(Surv(time, status) ~ ph.ecog.females, data = lung))

plot(surv)
```

7. (3) Calculate the hazard ratios comparing each level to patients in category 1.
``` {r}
lung$ph.ecog.females <- relevel(lung$ph.ecog.females,ref="1")
hr <- coxph(Surv(time, status) ~ ph.ecog.females,data = lung)
exp(cbind(hr$coef,confint(hr)))
```

8. (3) Calculate the hazard ratio for a 1 unit change in this variable (e.g. going from 1 to 2 or 2 to 3).
``` {r}
hra <- coxph(Surv(time,status) ~ ph.ecog, data=lung)
exp(cbind(hra$coef,confint(hra)))
```

9. (10) Examine the association of the variables age, ph.karno, pat.karno, meal.cal and wt.loss with survival using univariable Cox models and comment on the significance of the association with each.
``` {r}
summary(oslog1 <- survreg(Surv(time, status) ~ age, dist="lognormal", data = lung))
```
