---
title: "qbs124_hw2_gibran"
author: "Gibran Erlangga"
date: "4/9/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Use mortgageROC.csv data from Example 5.5 of the book.

``` {r}
data <- read.csv('mortgageROC.csv')
head(data)
```

##  Question 1
(20 points). Plot the two empirical cdfs with family income on the log scale with actual numbers displayed in thousand dollars. Use rug, legend, and different colors.
``` {r}
data$FamilyIncomeLog = log(data$FamilyIncome)

#a = default, b = no default
a = sort(data$FamilyIncomeLog[data$Default=="yes"])
b = sort(data$FamilyIncomeLog[data$Default=="no"])
label_range = range(c(2.5:7))

# compare cdfs
dollar_val = c(25,50,75,100,150,200,300,410)
xlabel_log = log(dollar_val)
label_range = range(xlabel_log)
x=seq(from=label_range[1],to=label_range[2],length=279)
plot(b, pnorm(b,mean=mean(b),sd=sd(b)), xlim=label_range, 
     col=3, lwd=2, type='s', 
     main="empirical cdf between default and non-default",
     xlab='Family income, thousand dollars',
     xaxt = "n",
     ylab='Density distribution')
axis(side=1,at=xlabel_log,labels=as.character (dollar_val))
lines(a, pnorm(a,mean=mean(a),sd=sd(a)), col=4, lwd=2, type='s')
rug(a,ticksize=0.1, col=4)
rug(b,ticksize=0.05, col=3)
legend("topleft",c("Default 'yes'","Default 'no'"),
       col=c(4,3),lwd=c(4,3),lty=1,cex=0.75,bg="gray95")
```
##  Question 2
(20 points). Create an R animation where at left you show empirical cdfs from the previous task and the growing stepwise ROC curve at right (use type="s") as in cdf.dyn(job=3). Submit as a standalone gif file.
``` {r}
n=100
# a => income for non-defaulters
X = b
# a => income for defaulters
Y = a
XY=sort(c(X,Y))
X=sort(X);Y=sort(Y)
nX=length(X);nY=length(Y)
n=length(XY)
th=XY
niY=niX=rep(NA,n)
for(i in 1:n)
{
  ch=as.character(i)
  if(i<10) ch=paste("00",ch,sep="")
  if(i>=10 & i<100) ch=paste("0",ch,sep="")       
  jpeg(paste("cdf",ch,".jpg",sep=""),width=1200,height=600)
  par(mfrow=c(1,2),mar=c(4.5,4.5,3,1),cex.lab=1.75,cex.main=1.75)
  dollar_val = c(25,50,75,100,150,200,300,410)
   xlabel_log = log(dollar_val)
   label_range = range(xlabel_log)
   x=seq(from=label_range[1],to=label_range[2],length=279)
  plot(XY,XY,type="n",ylim=c(0,1),
       xlab="Family income, thousand dollars",
       xaxt = "n",
       ylab="Proportion",
       main="Two CDFs for uniform data comparison: Y < X")
  axis(side=1,at=xlabel_log,labels=as.character (dollar_val))
  legend("topleft",c("Data","Threshold","cdf of default yes","cdf of default no"),col=c(1,1,3,2),lwd=c(1,3,3,3),cex=1.5,bg="gray95")
  rug(X,ticksize=0.075,col=2)
  rug(Y,ticksize=0.05,col=3)
  segments(th[i],-1,th[i],.05,lwd=3)
  segments(min(XY)-1,0,th[i],0)
  Xi=c(X[X<=th[i]],th[i])
  niX[i]=length(Xi)
  lines(Xi,(1:niX[i])/nX,type="s",col=2,lwd=2)
  Yi=c(Y[Y<=th[i]],th[i])
  niY[i]=length(Yi)
  lines(Yi,(1:niY[i])/nY,type="s",col=3,lwd=2)
  
  plot(niX[1:i]/nX,niY[1:i]/nY,xlim=c(0,1),ylim=c(0,1),lwd=3,type="s",xlab="1-Specificity (false positive)",ylab="Sensitivity (true positive)",main="ROC curve")			
  segments(-1,-1,2,2,col=4)
  
  dev.off()		
}

#I couldn't figure out magick in mac so I made the gif file from online source (ezgif.com)
```

##  Question 3
(20 points). (a) Compute AUC and provide its layman interpretation. 
``` {r}
library(pROC)

data$default_flag<-ifelse(data$Default=="yes",1,0)
model<-glm(default_flag~FamilyIncome, family="binomial", data=data)
predicted <- predict(model, data, type="response")

paste("AUC score:", round(auc(data$default_flag, predicted), 4))
```

AUC is the probability of Y < X, where Y represents the positive instance and X represents the negative instance. The AUC score in this particular case is 0.822 / 82.2%, which is a relatively high AUC value. An AUC score of 82% indicates that the probability of Y ranking higher than X is 82%.

(b) Compute and provide a layman interpretation for True Positive, True negative, False Positive, and False negative when Sensitivity=0.8. 
``` {r}
n_income <- length(data$FamilyIncome)

# prep data
default<- data[,1]
income <- data[,2]

income0 <- sort(income[default=="no"])
n0 <- length(income0)
m0 <- mean(income0)
s0 <- sd(income0)
income1 <- sort(income[default=="yes"])
n1 <- length(income1)
m1 <- mean(income1)
s1 <- sd(income1)
income<- sort(income)

sens=fp=toter10=rep(NA,n)
AUC=toter=0
for (i in 1:n) {
  sens[i] <- sum(income1 < income[i])/n1
  fp[i] <- sum(income0 < income[i])/n0
  toter10[i] <- (4.5/10)*(1-sens[i])+(5.5/10)*fp[i]
  if(i>1) AUC <- AUC+(fp[i]-fp[i-1])*sens[i]
}
opt.thresh <- unique(income[which(toter10==min(toter10))])
fp10 <- sum(income0<opt.thresh)/n0
plot(fp,sens,type="s",lwd=3, col=8,
     main="ROC curve for identification of default mortgages",
     xlab="False positive",
     ylab="Sensitivity")
abline(h = 0.8)
AUC.th <- pnorm((m1-m0)/sqrt(s0^2+s1^2))
segments(opt.thresh,-1,opt.thresh,2,col=2)
text(.72,.4, paste("AUC = ",round(100*AUC),"%"),cex=1.25,font=2)
```

``` {r}
# index position where sensitivity = .8, approximately
idx_80_sens <- 173

tp=sens[idx_80_sens]*n1
paste("The true positive is",tp)
fp_number=fp[idx_80_sens]*n0
paste("The false positive is",fp_number)
fn=n1-tp
paste("The false negative is",fn)
tn=n0-fp_number
paste("The true negative is",tn)
```
Sensitivity is showing the proportion of true positives over all actual positive cases. In this particular case, when sensitivity = 0.8, the value for true positive (TP), false positive (FP), false negative (FN), and true negative (TN) are 77, 95, 19, and 184, respectively.


(c) Compute and display the optimal threshold if the cost of overlooking a future defaulter is \$200K and the cost of denying the mortgage application who will not default in the future is \$100K.
``` {r}
sens=fp=toter10=rep(NA,n)
AUC=toter=0
for (i in 1:n) {
  sens[i] <- sum(income1 < income[i])/n1
  fp[i] <- sum(income0 < income[i])/n0
  # n1 = defaulters; n0 = non-defaulters
  toter10[i] <- (n1*200000)*(1-sens[i])+(n0*100000)*fp[i]
  if(i>1) AUC <- AUC+(fp[i]-fp[i-1])*sens[i]
}
opt.thresh <- unique(income[which(toter10==min(toter10))])
fp10 <- sum(income0<opt.thresh)/n0
plot(fp,sens,type="s",lwd=3, col=8,
     main="ROC curve for identification of the normal patient",
     xlab="False positive",
     ylab="Sensitivity")
segments(fp10,-1,fp10,2,col=2)
text(fp10+.01,1,"Optimal threshold",col=2,adj=0)
segments(opt.thresh,-1,opt.thresh,2,col=2)
AUC.th <- pnorm((m1-m0)/sqrt(s0^2+s1^2))
text(.72,.4,paste("AUC = ",round(100*AUC),
                  "%\nOptimal threshold =",
                  round(opt.thresh),sep=""),
                  cex=1.25,font=2)
```

